---
title: 'Football '
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Loading required libraries
```{r warning=FALSE, message=FALSE}
library(Hmisc)
library(caTools)
library(randomForest)
library(ggplot2)
library(plotly)
library(e1071)
library(ROCR)
library(pROC)
```
### Reading and summarizaiton of data
```{r}
data <- read.csv("results.csv") 
data<-data[(data$Year>="2008"),]
summary(data)
```
### How Data Looks:

```{r}
head(data)
```
```{r}
#Checking Missing Values

missing <- as.data.frame(apply(data, 2, FUN = function(x) sum(is.na(x))))
colnames(missing) <- "Missing_Count"
missing$Missing_Per <- (missing$Missing_Count/(nrow(data))*100)
missing$Missing_Per <- round(missing$Missing_Per,0)
missing
```
### Exploratory Analysis

#### Top 10 Match Host
```{r}
# Top ten match hosts 

host<-as.data.frame(tail(sort(table(data$country)),10))
colnames(host) <- c("Country", "No_of_Matches")

# Draw plot
graph_1<-ggplot(host, aes(x=Country,y=No_of_Matches)) + 
  geom_bar(stat="identity", width=.75, fill="Red") + 
  labs(title="Top 10 Host", 
       subtitle="2008-2018", 
       caption="source: Kaggle") 
graph_1
```

#### Home Advantage?
```{r}

home_win <- (as.character(data$country) == as.character(data$home_team)) & (as.integer(data$home_score) >= as.integer(data$away_score))  
away_win <- (as.character(data$country) == as.character(data$home_team)) & (as.integer(data$home_score) < as.integer(data$away_score))  

home_win<-as.data.frame(table(home_win))
away_win<-as.data.frame(table(away_win))


graph_2_a <- plot_ly(home_win, x = ~home_win, y = ~Freq, type = 'bar',
             marker = list(color = c('rgba(204,204,204,1)', 'rgba(222,45,38,0.8)'))) %>%
  layout(title = "Home Wins",
         yaxis = list(title = "# of Matches"))

graph_2_b <- plot_ly(away_win, x = ~away_win, y = ~Freq, type = 'bar',
             marker = list(color = c('rgba(204,204,204,1)', 'rgba(222,45,38,0.8)'))) %>%
  layout(title = "Away Wins",
         yaxis = list(title = "# of Matches"))
graph_2_a 
graph_2_b
```

#BEST TEAMS OF ALL THE TIME
```{r}
best_team <- (0)
for (i in 1:length(data$home_team))
  {
    x <- ifelse(data[i,4,] >= data[i,5,], as.character(data[i,2,]), as.character(data[i,3,]));
    best_team <- c(best_team, x)
  }

best_team<-as.data.frame(tail(sort(table(best_team)),10))

graph_3<-ggplot(best_team, aes(x=best_team,y=Freq)) + 
  geom_bar(stat="identity", width=.75, fill="Blue") + 
  coord_flip() +
  labs(title="Top 10 Home Teams", 
       subtitle="2008-2018", 
       caption="source: Kaggle") 
graph_3
```

####  Top 10 TOURNAMENTS 

```{r}
tournament<-as.data.frame(tail(sort(table(data$tournament)),10))
colnames(tournament) <- c("Tournament", "No_of_Matches")


graph_4<-plot_ly(tournament, x=~Tournament, y=~No_of_Matches, type = 'bar') %>% layout(title = "Top Tournaments")
graph_4

```

#### Brazil trend of playing matches over the years
```{r}

temp<-data.frame(data$home_team,data$away_team,data$Year)
team<-"Brazil"
colnames(temp) <- c("home_team", "away_team","year")
temp <- temp[temp$home_team == team | temp$away_team ==team,]
team_trend<-as.data.frame(tail(sort(table(temp$year)),10))
colnames(team_trend) <- c("year", "freq")
team_trend$year<-as.character(team_trend$year)

graph_5<-ggplot(data=team_trend, aes(x=year, y=freq, group=1)) +
  geom_line(colour="blue", linetype="dashed", size=1.5) + 
  geom_point(colour="blue", size=4, shape=21, fill="white")
graph_5

```


### Reading data for top Teams, Stadiums and Tournaments
```{r}
data <- read.csv("results.csv", stringsAsFactors = FALSE) 
home<-read.csv("best_teams.csv")
country<-read.csv("country.csv")
tournament<-read.csv("tournament.csv")
```
## Feature Engineering

#### Creating a winner variable

```{r}
for(i in 1:length(data$home_team))
{
  if(data$home_score[i]>data$away_score[i])
  {
    data$Winner[i]<-"Home"
  }
  else if(data$home_score[i]<data$away_score[i])
  {
    data$Winner[i]<-"Away"
  }
  else
  {
    data$Winner[i]<-"Tie"
  }
}
head(data)
```

#### Binning of Home and Away Team
```{r}

for ( i in 1:length(data$home_team))
{
  if (data$home_team[i] %nin% home$Teams){
    data[i,"home_team"] <- "others"
  }
}

for ( i in 1:length(data$away_team))
{
  if (data$away_team[i] %nin% home$Teams){
    data[i,"away_team"] <- "others"
  }
}
unique(data$home_team)
```

#### Binning of Tournaments 
```{r}
for ( i in 1:length(data$tournament))
{
  if (data$tournament[i] %nin% tournament$Tournament){
    data[i,"tournament"] <- "others"
  }
}
unique(data$tournament)
```

#### Removal of matches( others vs others) and less important tournaments
```{r}
data<-data[!(data$home_team=="others" & data$away_team=="others"),]
#24564
data<-data[!(data$tournament=="others"),]
#23911
```

#### Setting "Home" as country where Neutral is FALSE
```{r}
for(i in 1:length(data$neutral))
{
  if(data$neutral[i]=="FALSE")
  {
    data$country[i]<-"HOME"
  }
}
```

#### Binning of Country (Excluding the Home teams)
```{r}
for(i in 1:length(data$neutral))
{
  if(data$neutral[i]=="TRUE")
  {
    if (data$country[i] %nin% country$Country){
      data[i,"country"] <- "others"
    }
  }
}
unique(data$country)
```

#### Removing unnecessary columns not required for model
```{r}
#Date
data<-data[-1]
#city
data<-data[-6]
#Score
data<-data[-3:-4]
#Year
data<-data[-6]
```

```{r}
sapply(data, function(x) length(unique(x)))
sapply(data, class)
```

#### converting variables into factor 
```{r}
data$home_team<-as.factor(data$home_team)
data$away_team<-as.factor(data$away_team)
data$tournament<-as.factor(data$tournament)
data$country<-as.factor(data$country)
data$Winner<-as.factor(data$Winner)
data$neutral<-as.factor(data$neutral)
data$Month<-as.factor(data$Month)
```
## Model Creation
#### Splitting the data into Train and Test
```{r}
set.seed(123)
split = sample.split(data$Winner, SplitRatio = 0.75)
train_set = subset(data, split == TRUE)
test_set = subset(data, split == FALSE)
```
## Random Forest
#### Training the model
```{r}
set.seed(123)
classifier = randomForest(x = train_set[-7],
                          y = train_set$Winner,
                          ntree = 500)
```

#### Testing the trained model
```{r}
y_pred = predict(classifier, newdata = test_set[-7])
cm = table(test_set[, 7], y_pred)
```

#### calculating accuracy
```{r}
n = sum(cm) # number of instances
diag = diag(cm) # number of correctly classified instances per class
accuracy = sum(diag) / n
accuracy*100
```

## Naive Bayes
#### Training the model
```{r}
classifier = naiveBayes(x = train_set[-7],
                        y = train_set$Winner)
```

#### Testing the model
```{r}
y_pred = predict(classifier, newdata = test_set[-7])
cm = table(test_set[, 7], y_pred)
```

#### calculating accuracy
```{r}
n = sum(cm) # number of instances
diag = diag(cm) # number of correctly classified instances per class
accuracy = sum(diag) / n
accuracy*100
```

#### ROC
```{r}
y_pred<-as.data.frame(y_pred)
 roc.home <- roc(ifelse(test_set$Winner=="Home", "Home", "non-Home"), as.numeric(y_pred$y_pred))
 roc.away <- roc(ifelse(test_set$Winner=="Away", "Away", "non-Away"), as.numeric(y_pred$y_pred))
 roc.tie <- roc(ifelse(test_set$Winner=="Tie", "Tie", "non-Tie"), as.numeric(y_pred$y_pred))
 
 plot(roc.home, col = "green", main="ROC Curve")
 lines(roc.away, col = "blue")
 lines(roc.tie, col = "red")
 legend("topleft", c("Home","Away","Tie"), fill=c("green","blue","red") )
```

#### AUC
```{r}
 auc(roc.home)
 auc(roc.away)
 auc(roc.tie)
```
